---
title: "The Effects of Corruption on Monetary Wealth [Code, Output, Notes]"
author: "Patrick Poleshuk / Katherine Sprowls / Justin Kerner / Riley McCoy"
date: "12/06/2020"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
error_hook <- knitr::knit_hooks$get("error")
knitr::knit_hooks$set(error = function(x, options) {
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  error_hook(x, options)
})
```

```{r setup/analysis, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
library(ggplot2)
library(tidyverse)
library(WDI)
library(popbio)
library(dplyr)
library(plm)
library(tidyr)
library(car)

file = "/Users/patrickpoleshuk/Desktop/archive (1)/history.csv"
df_hist = read.csv(file, header = F, sep = ",")
df_hist = subset(df_hist, select = -c(V1, V3, V4))
names(df_hist) = c("Country", "Corruption - 2016", 
                   "Corruption - 2015", "Corruption - 2014", 
                   "Corruption - 2013", "Corruption - 2012")

df_hist <- df_hist %>% slice(-c(1))

df_literacy = WDI(country = "all", indicator = "SE.ADT.LITR.ZS", start = "2012",
                   end = "2016")
# Measures the literacy rate of the % of people, within a given country, 
# aged 15 and above. 

df_primary_completion = WDI(country = "all", indicator = "SE.PRM.CMPT.ZS",
                 start = "2012", end = "2016")
# Measures the primary school completion rate, as a % of the relevant age 
# group population, within a country. 

df_allocation = WDI(country = "all", indicator = "SE.XPD.TOTL.GD.ZS", start = "2012",
                    end = "2016")
# Measures the amount of revenue (as % of a country's GDP) that gets allocated towards education. 

df = merge(df_literacy, df_primary_completion)
df = merge(df_allocation, df)

names(df) <- c("code", "Country", "Year", "%_of_GDP_To_Education",  "Literacy_Rate",
               "%_Primary_Complete")
df = subset(df, select = -c(code))


df2 = WDI(country = "all", indicator = "NY.GDP.MKTP.CD", start = "2012",
         end = "2016")


keycol <- "condition"
valuecol <- "measurement"
names(df_hist) = c("Country", "C16", "C15", "C14", "C13", "C12" )
gathercols <- c("C16", "C15", "C14", "C13", "C12")

df_hist = gather_(df_hist, keycol, valuecol, gathercols)

df_hist <- df_hist[order(df_hist$Country, df_hist$condition), ]

attach(df_hist)
df_hist$Year[condition == "C12"] <- "2012"
df_hist$Year[condition == "C13"] <- "2013"
df_hist$Year[condition == "C14"] <- "2014"
df_hist$Year[condition == "C15"] <- "2015"
df_hist$Year[condition == "C16"] <- "2016"
detach(df_hist)

df_hist = subset(df_hist, select = -condition)

df_complete = (merge(df_hist, df, by=c("Country", "Year")))

NROW(df_complete)
View(df_complete)
names(df_complete) = c("Country", "Year", "Corruption_Index", "%_of_GDP_To_Education",
                       "Literacy_Rate", "%_Primary_Complete")

df2 = subset(df2, select = -c(iso2c))
names(df2) <- c("Country", "GDP", "Year")
df_complete = merge(df_complete, df2, by = c("Country", "Year"))

lnGDP = log(df_complete$GDP)

Corruption = as.numeric(df_complete$Corruption_Index)
Education_Allocation = df_complete$`%_of_GDP_To_Education`
Lit = df_complete$Literacy_Rate
Completion = df_complete$`%_Primary_Complete`

reg <- function(y, x){
  results = glm(y ~ x)
  return(results)
}

# * I'm wrapping the regression command into a function so it resembles more of something that is 
# seen in Stata. When you see "glm" or "plm" later on it's just a regression command.

cor.test(lnGDP, Corruption)
# A less corrupt country is moderately, postively correlated with a higher GDP at .4046. 
library(outreg)
summary(reg(lnGDP, Corruption))
reg(lnGDP, Corruption)
coef(reg(lnGDP, Corruption))[2] * 100
# A 1 unit increase in the corruption index will lead to a 4.1765% increase in GDP. 
# Statistically Significant (S.S) at the 1% signficance level. 
summary(reg(Education_Allocation, Corruption))
m <- reg(Education_Allocation, Corruption)
m
library(outreg)
outreg(m)
coef(reg(Education_Allocation, Corruption))[2]
# The higher the corruption index value, the least corrupt the country is. That's why the coefficent 
# of effect is positive. For every 1 index point increase in Corruption, there will be
# a .03752 increase in education allocated as a % of a country's total GDP. 
summary(reg(Lit, Corruption))
# A 1 unit increase in the corruption index is predicted to raise literacy rates in a
# country by an estimated .3595%. It should be noted that all our findings will be
# S.S. at the 1% level if not stated. 
reg(Completion, Corruption)
# A 1 unit increase in the corruption index will raise the primary school completion rate by .3432%. 

cor.test(Education_Allocation, Corruption)
# We found a correlation coefficent of .4966988. This signals a moderate linear relationship between
# corruption in a country and the % of GDP Allocated towards Education. 
length(which(!is.na(Education_Allocation & Corruption)))
# N = 454
freqData <- as.data.frame(table(Corruption, Education_Allocation))
names(freqData) <- c("Corruption", "Education", "Frequency")
freqData$Corruption <- as.numeric(as.character(freqData$Corruption))
freqData$Education <- as.numeric(as.character(freqData$Education))
g <- ggplot(filter(freqData, Frequency > 0), aes(x = Corruption, y = Education))
g <- g + scale_size(range = c(1, 8), guide = "none" )
g <- g + geom_point(color="grey50", aes(size = Frequency+20))
g <- g + geom_point(aes(color=Frequency, size = Frequency))
g <- g + scale_color_gradient(low = "lightblue", high="white") +
  ggtitle("Illustrated Regression of Corruption Index on Education Allocation") +
  labs(x = "Corruption Index",
       y = "Education Allocation (% of GDP)",
       caption = "Data Source: World Bank / Transparency International",
       subtitle = "N = 454")
Model_new <- glm(Education_Allocation ~ Corruption)
g <- g + geom_abline(intercept = coef(Model_new)[1], slope = coef(Model_new)[2], 
                     size = 3, color = "blue")
g
# Ignore the confusing syntax, this is just a graph of how the corruption index of a country
# predicts a higher % of GDP allocted to education. You'll see that the index of % allocated to
# education only goes up to about 10, that's because no country is willing to allocate more than 
# 10% of their respective GDP value, in resources, to their education sector. 

df_GDP_capita = WDI(country = "all", indicator = "NY.GDP.PCAP.CD", 
                    start = "2012", end = "2019")
df_GDP_capita <- subset(df_GDP_capita, select = c(country, NY.GDP.PCAP.CD, year))
names(df_GDP_capita) <- c("Country", "GDP_per_capita", "Year")

summary(Corruption)
attach(df_complete)
df_complete$Code[Corruption < 38] <- "Upper half of Corruption"
df_complete$Code[Corruption >= 38] <- "Lower Half of Corruption"
detach(df_complete)

CorruptionCode<- ifelse(df_complete$Code == "Upper half of Corruption", 0, 1)

df_complete = merge(df_complete, df_GDP_capita, by = c("Country", "Year"))
GDP_capita = df_complete$GDP_per_capita
lnGDP_capita = log(df_complete$GDP_per_capita)

length(which(!is.na(lnGDP_capita & CorruptionCode)))
# N = 745
probit <- ggplot(df_complete, aes(x=lnGDP_capita, y=CorruptionCode)) + geom_point() +
  stat_smooth(method = "glm", method.args = list(family = "binomial" (link = "probit")), se=TRUE)
probit + labs(x="Ln(GDP Per Capita)", y="Corruption (0 - Most, 1 - Least)", 
          title = "Probit Regression of Corruption on Log Value of GDP Per Capita", 
          caption = "Data Source: World Bank / Transparency International",
          subtitle = "N = 719 (Removed 26 Non-Finite Values)")
# Here I just changed the GDP_per_capita value into a log transformed variable, the "S" shape 
# is what we desire from the model. It is clear that more corrupt regimes are characterized by
# weaker GDP per capita values. 

summary(glm(GDP_capita ~ CorruptionCode))
# A country with corruption values below the median will tend to have a GDP per capita
# $20,732.8 less than countries who have a corruption value above the median. 
summary(reg(Education_Allocation, CorruptionCode))
# A country whose corruption index lies above the median allocated roughly 1.2% more to their 
# education sector than countries below the median. 
Corruption = as.numeric(df_complete$Corruption_Index)

cor.test(log(GDP_capita), Corruption)
# We see that GDP per capita is strongly linearly related with a lower corruption at a
# correlation coefficent of .7829. 

```

## Productivity Analysis

```{r Effect of Productivity, warning=FALSE}
file = "/Users/patrickpoleshuk/Downloads/DP_LIVE_09112020234158064.csv"
OECD = read.csv(file)

names(OECD) <- c("Country_Code", "r1", "r2", "r3", "r4", "Year", "Productivity", "r5")

OECD = subset(OECD, select = -c(r1, r2, r3, r4, r5))

df_literacy = WDI(country = "all", indicator = "SE.ADT.LITR.ZS", start = "1960",
                   end = "2019", extra = TRUE)

df_primary_completion = WDI(country = "all", indicator = "SE.PRM.CMPT.ZS",
                 start = "1960", end = "2019", extra = TRUE)

df_allocation = WDI(country = "all", indicator = "SE.XPD.TOTL.GD.ZS", start = "1960",
                    end = "2019", extra = TRUE)

df_GDP_capita = WDI(country = "all", indicator = "NY.GDP.PCAP.CD", 
                    start = "1960", end = "2019", extra = TRUE)

df = merge(df_literacy, df_primary_completion)
df = merge(df_allocation, df)
df = merge(df, df_GDP_capita)
df = subset(df, select = c(year, iso3c, SE.XPD.TOTL.GD.ZS, 
                           SE.PRM.CMPT.ZS, SE.ADT.LITR.ZS, NY.GDP.PCAP.CD))

names(df) <- c("Year", "Country_Code", "%_of_GDP_To_Education", "%_Primary_Complete",
              "Literacy_Rate", "GDP_Per_Capita")

df_final = merge(df, OECD, by = c("Country_Code", "Year"))

Productivity = df_final$Productivity
Edu_Allocation = df_final$`%_of_GDP_To_Education`
Literacy = df_final$Literacy_Rate

summary(Edu_Allocation)
attach(df_final)
df_final$Test[Edu_Allocation < 5.124] <- "Least Allocated"
df_final$Test[Edu_Allocation >= 5.124] <- "Most Allocated"
detach(df_final)

AllocationCode<- ifelse(df_final$Test == "Least Allocated", 0, 1)

summary(glm(Productivity ~ AllocationCode))
# Countries with an above average (median) allocation (as a % of total GDP) towards Education will 
# experience a 6.7542 increase in labor productivity (GDP per hour worked); significant 
# at the 1% level. 
summary(glm(Productivity ~ Edu_Allocation))
# A 1 % increase in Education allocated (as a total % of GDP will increase total productivity 
# (GDP per labor hour worked) in a given country by 4.6679. In other words, labor will produce
# 4.6679 dollars more of GDP per hour worked with greater allocation in the education sector.
summary(glm(Productivity ~ Literacy))
# A 1% increase in Literacy Rates for the net population will increase GDP per hour worked, 
# for the labor in that country, by 1.0423.

summary(glm(Productivity ~ df_final$`%_Primary_Complete`))
# A 1% increase in primary school completion rate increases GDP per hour worked by .0995. 
# However, this is not signficant at the 1% level nor the 10% level. 

p = df_final$Productivity
Allo = df_final$`%_of_GDP_To_Education`

length(which(!is.na(df_final$Productivity & df_final$`%_of_GDP_To_Education`)))
# N = 2,284
freqData <- as.data.frame(table(Allo, p))
names(freqData) <- c("Allo", "p", "Frequency")
freqData$Allo <- as.numeric(as.character(freqData$Allo))
freqData$p <- as.numeric(as.character(freqData$p))
g <- ggplot(filter(freqData, Frequency > 0), aes(x = Allo, y = p))
g <- g  + scale_size(range = c(1, 8), guide = "none" )
g <- g + geom_point(color="grey50", aes(size = Frequency+20))
g <- g + geom_point(aes(color=Frequency, size = Frequency))
g <- g + scale_color_gradient(low = "lightblue", high="white") +
  ggtitle("Illustrated Regression of Education Allocation on Worker Productivity") +
  labs(x = "Education Allocation (% of GDP)",
       y = "GDP Per Hour Worked (Productivity Index)",
       caption = "Data Source: World Bank / OECD",
       subtitle = "N = 2,284")
Model_new <- glm(p ~ Allo)
g <- g + geom_abline(intercept = coef(Model_new)[1], slope = coef(Model_new)[2], 
                     size = 3, color = "dark red")
g 

# Here we see the regression, stating that a greater education allocation is predicted to 
# increase labor productivity. 

GDP = df_final$GDP_Per_Capita

lnGDP = log(GDP)
Literacy = df_final$Literacy_Rate

fixed_effects = plm(lnGDP ~ Productivity + factor(Year), index = "Country_Code", 
                    data = df_final, model = "within")
summary(fixed_effects)
# We get a very strong r^2 value, .92907. This signifies that a variance in productivity can 
# explain roughly 93% of the variance in GDP per capita. We also get more precise estimators, 
# as in fixed effect models we control for lagged effects or the possibility of unobserved 
# heterogenity. On the top of our factor(year) variables we get a Productivity estimator 
# of .001253. For every 1$ increase in GDP per hour worked, we see a 1.253% increase in GDP 
# per capita. This is going to be our most refined estimator. This is S.S. at the 1% level, 
# but you will see our other fixed effect estimator will not be, and is, thus, not reliable. 

fixed_effects = plm(p ~ Allo + factor(Year), index = "Country_Code", 
                    data = df_final, model = "within")
summary(fixed_effects)
summary(glm(p~Allo)) # A 1 increase in education allocation increases productivity by 4.6679.
# A variance in Education Allocation explains roughly 25.48% of variance in productivity. 
# This is not surprising, as GDP produced per hour of labor is dependent on many variables such 
# as general company managament, access to technology, and profititability incentives.
fe = plm(p ~ Allo + factor(Country_Code), index = "Year", 
                    data = df_final, model = "within")
summary(fe)
# Controlling for aggregate trends that vary throughout country, in another fixed-effects
# model, 
# we find that an increase in education allocation as a % of GDP increases GDP per hour
# worked by 5.04463. 
# This is statistically significant at the 1% level, however, we lack of a lot of data on individual 
# countries to give a statistically signficant estimator for each country code category.
