---
title: "The Effects of Corruption on Monetary Wealth"
author: "Patrick Poleshuk / Katherine Sprowls / Justin Kerner / Riley McCoy"
date: "11/16/2020"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
error_hook <- knitr::knit_hooks$get("error")
knitr::knit_hooks$set(error = function(x, options) {
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  error_hook(x, options)
})
```

```{r setup/analysis, include=TRUE, message=FALSE, warning=FALSE}
# The goal of the below research is to point out that corruption is deleterious to the monetary 
# welfare of a country's citizens. We will examine this through interpreting statistical trends of
# the misallocation of scarce resources with alternative uses, as it pertains to the distribution of
# a country's GDP into its health and education sectors. Overall we find that corrupt regimes are
# more likely to allocate a lower amount of resources, relative to their GDP, into these sectors,
# and suffer from weaker labor productivity as a result. As we see, when labor is less productive
# GDP per capita of a country falls, and, on a macroeconomic scale, the entire economy suffers. 
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
library(ggplot2)
library(tidyverse)
library(WDI)
library(popbio)
library(dplyr)
library(plm)
library(tidyr)
library(car)
library(OECD)


file = "/Users/patrickpoleshuk/Desktop/archive (1)/history.csv"
df_hist = read.csv(file, header = F, sep = ",")
df_hist = subset(df_hist, select = -c(V1, V3, V4))
names(df_hist) = c("Country", "Corruption - 2016", 
                   "Corruption - 2015", "Corruption - 2014", 
                   "Corruption - 2013", "Corruption - 2012")

df_hist <- df_hist %>% slice(-c(1))

df_literacy = WDI(country = "all", indicator = "SE.ADT.LITR.ZS", start = "2012",
                   end = "2016")
# Measures the literacy rate of the % of people, within a given country, aged 15 and above. 

df_primary_completion = WDI(country = "all", indicator = "SE.PRM.CMPT.ZS",
                 start = "2012", end = "2016")
# Measures the primary school completion rate, as a % of the relevant age group population, within 
# a country. 

df_allocation = WDI(country = "all", indicator = "SE.XPD.TOTL.GD.ZS", start = "2012",
                    end = "2016")
# Measures the amount of revenue (as % of a country's GDP) that gets allocated towards education. 

df = merge(df_literacy, df_primary_completion)
df = merge(df_allocation, df)

names(df) <- c("code", "Country", "Year", "%_of_GDP_To_Education",  "Literacy_Rate", "%_Primary_Complete")
df = subset(df, select = -c(code))


df2 = WDI(country = "all", indicator = "NY.GDP.MKTP.CD", start = "2012",
         end = "2016")


keycol <- "condition"
valuecol <- "measurement"
names(df_hist) = c("Country", "C16", "C15", "C14", "C13", "C12" )
gathercols <- c("C16", "C15", "C14", "C13", "C12")

df_hist = gather_(df_hist, keycol, valuecol, gathercols)

df_hist <- df_hist[order(df_hist$Country, df_hist$condition), ]

attach(df_hist)
df_hist$Year[condition == "C12"] <- "2012"
df_hist$Year[condition == "C13"] <- "2013"
df_hist$Year[condition == "C14"] <- "2014"
df_hist$Year[condition == "C15"] <- "2015"
df_hist$Year[condition == "C16"] <- "2016"
detach(df_hist)

df_hist = subset(df_hist, select = -condition)

df_complete = (merge(df_hist, df, by=c("Country", "Year")))

NROW(df_complete)
View(df_complete)
names(df_complete) = c("Country", "Year", "Corruption_Index", "%_of_GDP_To_Education", "Literacy_Rate", "%_Primary_Complete")

df2 = subset(df2, select = -c(iso2c))
names(df2) <- c("Country", "GDP", "Year")
df_complete = merge(df_complete, df2, by = c("Country", "Year"))

lnGDP = log(df_complete$GDP)

Corruption = as.numeric(df_complete$Corruption_Index)
Education_Allocation = df_complete$`%_of_GDP_To_Education`
Lit = df_complete$Literacy_Rate
Completion = df_complete$`%_Primary_Complete`

reg <- function(y, x){
  results = glm(y ~ x)
  return(results)
}

# * I'm just wrapping the regression command into a function so it resembles more of something that 
# you guys have seen in Stata. When you see "glm" or "plm" later on it's just a regression command.

cor.test(lnGDP, Corruption)
# A less corrupt country is moderately, postively correlated with a higher GDP at .4046. 
# Not surprising, as there are a myriad of factors that impact GDP. 
summary(reg(lnGDP, Corruption))
# A 1 unit increase in the corruption index will lead to a 041765% increase in GDP. S.S. at 1% signficant level. 

summary(reg(Education_Allocation, Corruption))
# Remember! In this data frame, the higher the Corruption index value the least corrupt 
# the country is. That's why the coefficent of effect is positive. Essentially, 
# For every 1 index point increase in Corruption, there will be a .03752% increase 
# in education allocated as a % of a country's total GDP. 
summary(reg(Lit, Corruption))
# A 1 unit increase in the corruption index is predicted to raise literacy rates in that country 
# by an estimated .3595%. 
m = glm(Lit ~ Corruption + Education_Allocation)
print(m)
# A quick multiple linear regression reveals these findings which you can interpret. 
vif(m)
# If the Variance inflation factor is even close to 1, there are no issues with multicollinearity. 
# So we're fine. 
summary(m)
# All variables are signficant at the 1% level. And it will go without saying later on, 
# that all our findings will be S.S. at at least the 5% level unless I say so. 
reg(Completion, Corruption)
# A 1 unit increase in the corruption index will raise the primary school completion rate by .3432%. 
model = glm(Lit ~ Corruption + Education_Allocation + Completion)
summary(model)
# In this multiple linear regression, we can see that all of these variables 
# have a great impact on literacy rates. 
# Problem is that this is the only model with a weak S.S. of the variables. 
vif(model)
# Doubt it's an issue with multicollinearity, probably just a sample size error. 

cor.test(Education_Allocation, Corruption)
# We found a correlation coefficent of .4966988. This signals a moderate linear relationship between
# corruption in a country and the % of GDP Allocated towards Education. 
freqData <- as.data.frame(table(Corruption, Education_Allocation))
names(freqData) <- c("Corruption", "Education", "Frequency")
freqData$Corruption <- as.numeric(as.character(freqData$Corruption))
freqData$Education <- as.numeric(as.character(freqData$Education))
g <- ggplot(filter(freqData, Frequency > 0), aes(x = Corruption, y = Education))
g <- g  + scale_size(range = c(1, 8), guide = "none" )
g <- g + geom_point(color="grey50", aes(size = Frequency+20))
g <- g + geom_point(aes(color=Frequency, size = Frequency))
g <- g + scale_color_gradient(low = "lightblue", high="white") +
  ggtitle("Illustrated Regression of Corruption Index on GDP") +
  labs(x = "Corruption Index",
       y = "Education Allocation")
Model_new <- glm(Education_Allocation ~ Corruption)
g <- g + geom_abline(intercept = coef(Model_new)[1], slope = coef(Model_new)[2], size = 3, color = "blue")
g

# Ignore the confusing syntax, this is just a nice graph of how the corruption index of a country
# predicts a higher % of GDP allocted to education. You'll see that the index of % allocated to
# education only goes up to about 10, that's because no country is willing to allocate more than 
# 10% of their respective GDP value, in resources, to their education sector. 

plot(Corruption, Education_Allocation, main = "Education Allocation Alongside Corruption Index Plot",
     xlab = "Corruption Index", ylab = "Education Allocation as a % of GDP")
abline(glm(Education_Allocation ~ Corruption), col = "Blue")
# This is watered down, simple graph of the same findings. 

df_GDP_capita = WDI(country = "all", indicator = "NY.GDP.PCAP.CD", 
                    start = "2012", end = "2019")
df_GDP_capita <- subset(df_GDP_capita, select = c(country, NY.GDP.PCAP.CD, year))
names(df_GDP_capita) <- c("Country", "GDP_per_capita", "Year")

summary(Corruption)
attach(df_complete)
df_complete$Code[Corruption < 38] <- "Upper half of Corruption"
df_complete$Code[Corruption >= 38] <- "Lower Half of Corruption"
detach(df_complete)

CorruptionCode<- ifelse(df_complete$Code == "Upper half of Corruption", 0, 1)

df_complete = merge(df_complete, df_GDP_capita, by = c("Country", "Year"))
GDP_capita = df_complete$GDP_per_capita
lnGDP_capita = log(df_complete$GDP_per_capita)

probit <- ggplot(df_complete, aes(x=GDP_capita, y=CorruptionCode)) + geom_point() +
  stat_smooth(method = "glm", method.args = list(family = "binomial" (link = "probit")), se=TRUE)
probit + labs(x="GDP per capita", y="Corruption (0 - Most, 1 - Least)", 
          title = "Probit Regression of Corruption on GDP per capita")

# Ok here is a probit (Non-linear regression) model of GDP per capita, depending on if a 
# country falls below or above the median corruption index value of 38. I will refine this in 
# the next line of code. 

probit <- ggplot(df_complete, aes(x=lnGDP_capita, y=CorruptionCode)) + geom_point() +
  stat_smooth(method = "glm", method.args = list(family = "binomial" (link = "probit")), se=TRUE)
probit + labs(x="Log valuee of GDP per capita", y="Corruption (0 - Most, 1 - Least)", 
          title = "Probit Regression of Corruption on log value of GDP per capita")


# Here I just changed the GDD_per_capita value into a log transformed variable, the "S" shape 
# is what we desire from the model. It is clear that more corrupt regimes are characterized by
# weaker GDP per capita values. 

summary(glm(GDP_capita ~ CorruptionCode))
# A country with corruption values below the median will tend to have a GDP per capita 
# $20,732.8 less than countries who have a corruption value above the median. 
df_health = WDI(country = "all", indicator = "SH.XPD.CHEX.GD.ZS", start = 1960, 
                end = 2019)

df_health = subset(df_health, select = -c(iso2c))
names(df_health) <- c("Country", "Health_Expenditure_%_of_GDP", "Year")

df_complete = merge(df_health, df_complete, by = c("Country", "Year"))
health = df_complete$`Health_Expenditure_%_of_GDP`
allocation = df_complete$`%_of_GDP_To_Education`
Corruption = as.numeric(df_complete$Corruption_Index)

summary(glm(health ~ Corruption))
# As the corruption index increases by 1, health expenditure as a % of GDP, as with education
# expenditure, increases by .0536. This is not surprising, as with greater governmental corruption
# we can presume that there exists a greater misallocation of resources.
```

## R Markdown

```{r Effect of Productivity, warning=FALSE}
file = "/Users/patrickpoleshuk/Downloads/DP_LIVE_09112020234158064.csv"
OECD = read.csv(file)

names(OECD) <- c("Country_Code", "r1", "r2", "r3", "r4", "Year", "Productivity", "r5")

OECD = subset(OECD, select = -c(r1, r2, r3, r4, r5))

df_literacy = WDI(country = "all", indicator = "SE.ADT.LITR.ZS", start = "1960",
                   end = "2019", extra = TRUE)

df_primary_completion = WDI(country = "all", indicator = "SE.PRM.CMPT.ZS",
                 start = "1960", end = "2019", extra = TRUE)

df_allocation = WDI(country = "all", indicator = "SE.XPD.TOTL.GD.ZS", start = "1960",
                    end = "2019", extra = TRUE)

df_GDP_capita = WDI(country = "all", indicator = "NY.GDP.PCAP.CD", 
                    start = "1960", end = "2019", extra = TRUE)

df = merge(df_literacy, df_primary_completion)
df = merge(df_allocation, df)
df = merge(df, df_GDP_capita)
df = subset(df, select = c(year, iso3c, SE.XPD.TOTL.GD.ZS, SE.PRM.CMPT.ZS, SE.ADT.LITR.ZS, NY.GDP.PCAP.CD))

names(df) <- c("Year", "Country_Code", "%_of_GDP_To_Education", "%_Primary_Complete", "Literacy_Rate", "GDP_Per_Capita")

df_final = merge(df, OECD, by = c("Country_Code", "Year"))

Productivity = df_final$Productivity
Edu_Allocation = df_final$`%_of_GDP_To_Education`
Literacy = df_final$Literacy_Rate

summary(Edu_Allocation)
attach(df_final)
df_final$Test[Edu_Allocation < 5.124] <- "Least Allocated"
df_final$Test[Edu_Allocation >= 5.124] <- "Most Allocated"
detach(df_final)

AllocationCode<- ifelse(df_final$Test == "Least Allocated", 0, 1)

summary(glm(Productivity ~ AllocationCode))
# Countries with an above average (median) allocation (as a % of total GDP) towards Education will 
# experience a 6.7542 increase in labor productivity (GDP per hour worked); significant 
# at the 1% level. 
summary(glm(Productivity ~ Edu_Allocation))
# A 1 % increase in Education allocated (as a total % of GDP will increase total productivity 
# (GDP per labor hour worked) in a given country by 4.6679. In other words, labor will produce
# 4.6679 dollars more of GDP per hour worked with greater allocation in the education sector.
summary(glm(Productivity ~ Literacy))
# A 1% increase in Literacy Rates for the net population will increase GDP per hour worked, 
# for the labor in that country, by 1.0423.


p = df_final$Productivity
Allo = df_final$`%_of_GDP_To_Education`

plot(Allo, p, main = "Worker Productivity Alongside Education Allocation",
     xlab = "Education Allocation as a % of GDP", ylab = "GDP per hour worked (Productivity Index)")
abline(glm(p ~ Allo), col = "Blue")

freqData <- as.data.frame(table(Allo, p))
names(freqData) <- c("Allo", "p", "Frequency")
freqData$Allo <- as.numeric(as.character(freqData$Allo))
freqData$p <- as.numeric(as.character(freqData$p))
g <- ggplot(filter(freqData, Frequency > 0), aes(x = Allo, y = p))
g <- g  + scale_size(range = c(1, 8), guide = "none" )
g <- g + geom_point(color="grey50", aes(size = Frequency+20))
g <- g + geom_point(aes(color=Frequency, size = Frequency))
g <- g + scale_color_gradient(low = "lightblue", high="white") +
  ggtitle("Illustrated Regression of Corruption Index on GDP") +
  labs(x = "GDP per hour worked (Productivity Index)",
       y = "Education Allocation as a % of GDP")
Model_new <- glm(p ~ Allo)
g <- g + geom_abline(intercept = coef(Model_new)[1], slope = coef(Model_new)[2], size = 3, color = "dark red")
g

# Here we see the regression, stating that a greater education allocation is predicted to 
# increase labor productivity. 

GDP = df_final$GDP_Per_Capita

lnGDP = log(GDP)
Literacy = df_final$Literacy_Rate
m <- glm(lnGDP ~ Productivity + Literacy)
summary(m)
# For every 1$ increase in GDP per hour worked, GDP per capita will go up by .5213% and for every 
# 1% increase in Literacy rates GDP per capita will increase by 6%. This may seem like an
# exaggeration but both these coefficent of effects are S.S. at the 1% level and if we refer back 
# to the probit model, the level of GDP per capita dispairty between corrupt and 
# non-corrupt countries was huge.
cor.test(lnGDP, Productivity)
# There is a .562843 correlation coefficent, indicating that there is a moderate linear relationship
# between higher productivity and GDP per capita. 

fixed_effects = plm(lnGDP ~ Productivity + factor(Year), index = "Country_Code", data = df_final, model = "within")
summary(fixed_effects)
# We get a very strong r^2 value, .92907. This signifies that a variance in productivity can 
# explain roughly 93% of the variance in GDP per capita. We also get more precise estimators, 
# as in fixed effect models we control for lagged effects or the possibility of reverse 
# casuality. Anyway, on the top of our factor(year) variables we get a Productivity estimator 
# of .001253. For every 1$ increase in GDP per hour worked, we see a 1.253% increase in GDP 
# per capita. This is going to be our most refined estimator. This is S.S. at the 1% level, 
# but you will see our other fixed effect estimators will not be, and are, thus, not reliable. 

fixed_effects = plm(p ~ Allo + factor(Year), index = "Country_Code", data = df_final, model = "within")
summary(fixed_effects)
# A variance in Education Allocation explains roughly 25.48% of variance in productivity. 
# This is not surprising, as GDP produced per hour of labor is dependent on many variables such 
# as general company managament, access to technology, and profititability incentives. Not a bad value.

df_health = WDI(country = "all", indicator = "SH.XPD.CHEX.GD.ZS", start = 1960, 
                end = 2019, extra = TRUE)
df_health = rename(df_health, "Health" = "SH.XPD.CHEX.GD.ZS")
library(sqldf)
df_health = sqldf("Select iso3c, year, Health from df_health")
names(df_health) <- c("Country_Code", "Year", "Health_%_of_GDP")

df_final = merge(df_final, df_health, by = c("Country_Code", "Year"))

fixed_effects = plm(df_final$Productivity ~ df_final$`Health_%_of_GDP` + factor(Year), index = "Country_Code", data = df_final, model = "within")
summary(fixed_effects)

# Unfortunately, we found a weak r^2 value of % allocated to GDP and productivity of a country's
# labor. Only roughly 4.74% of a variance in this variable explains a variance in productivity.
# However, not all these coefficients for the factor of years is S.S. at the 1% level, so maybe 
# we just lack the appropiate sample size for these years to make a truly informed interpretation. 

summary(glm(df_final$Productivity ~ df_final$`Health_%_of_GDP`))
# A 1% increase in resources (as a % of GDP) allocated towards health expenditures is predicted 
# to increase GDP per hour worked by 3.5794. 
cor.test(df_final$Productivity, df_final$`Health_%_of_GDP`)
# Correlation coefficent of .2726, fairly weak. 

df_health_exp_capita = WDI(country = "all", indicator = "SH.XPD.CHEX.PC.CD", start = 1960, end = 2019, extra = TRUE)

df_health_exp_capita = rename(df_health_exp_capita, "HealthCapita" = "SH.XPD.CHEX.PC.CD")

df_health_exp_capita = sqldf("select iso3c, HealthCapita, year from df_health_exp_capita")

names(df_health_exp_capita) <- c("Country_Code", "Health_Exp_per_capita", "Year")

df_final = merge(df_health_exp_capita, df_final, by = c("Country_Code", "Year"))

cor.test(log(df_final$Health_Exp_per_capita), df_final$Productivity)

fe = plm(df_final$Productivity ~ log(df_final$Health_Exp_per_capita) + factor(Year), index = "Country_Code", data = df_final, model = "within")
summary(fe)

# Honestly, fairly same results as before, I tried a new data set for health and roughly the same
# results came up. I used Health expenditures per capita instead of % of GDP allocated towards 
# the health sector of a given country's economy. 
summary(glm(df_final$Productivity ~ log(df_final$Health_Exp_per_capita)))
# For every 1% increase in Health Expenditures per capita, GDP per hour worked will increase 
# by 10.88. Extremely high, but keep in mind a 1% increase in health expenditure 
# per capita is no small feat. 